---
title: "week11 PCA"
author: "Yiheng Hu"
date: "July 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Principal Component Analysis in R
#Compute the Principal Components using prcomp function. Plotting PCA components and illustrate the result.
```{r, echo=FALSE}
#Principal Component Analysis in R
#Compute the Principal Components
mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE)



summary(mtcars.pca)

## Importance of components:
##                           PC1    PC2     PC3     PC4     PC5     PC6
## Standard deviation     2.3782 1.4429 0.71008 0.51481 0.42797 0.35184
## Proportion of Variance 0.6284 0.2313 0.05602 0.02945 0.02035 0.01375
## Cumulative Proportion  0.6284 0.8598 0.91581 0.94525 0.96560 0.97936
##                            PC7    PC8     PC9
## Standard deviation     0.32413 0.2419 0.14896
## Proportion of Variance 0.01167 0.0065 0.00247
## Cumulative Proportion  0.99103 0.9975 1.00000

##PC1 and PC2 can explain 86% of the variance

str(mtcars.pca)

## List of 5
##  $ sdev    : num [1:9] 2.378 1.443 0.71 0.515 0.428 ...
##  $ rotation: num [1:9, 1:9] -0.393 0.403 0.397 0.367 -0.312 ...
##   ..- attr(*, "dimnames")=List of 2
##   .. ..$ : chr [1:9] "mpg" "cyl" "disp" "hp" ...
##   .. ..$ : chr [1:9] "PC1" "PC2" "PC3" "PC4" ...
##  $ center  : Named num [1:9] 20.09 6.19 230.72 146.69 3.6 ...
##   ..- attr(*, "names")= chr [1:9] "mpg" "cyl" "disp" "hp" ...
##  $ scale   : Named num [1:9] 6.027 1.786 123.939 68.563 0.535 ...
##   ..- attr(*, "names")= chr [1:9] "mpg" "cyl" "disp" "hp" ...
##  $ x       : num [1:32, 1:9] -0.664 -0.637 -2.3 -0.215 1.587 ...
##   ..- attr(*, "dimnames")=List of 2
##   .. ..$ : chr [1:32] "Mazda RX4" "Mazda RX4 Wag" "Datsun 710" "Hornet 4 Drive" ...
##   .. ..$ : chr [1:9] "PC1" "PC2" "PC3" "PC4" ...
##  - attr(*, "class")= chr "prcomp"


#plotting PCA
library(ggbiplot)
ggbiplot(mtcars.pca)
ggbiplot(mtcars.pca, labels=rownames(mtcars))
str(mtcars)

#Interpreting the results
mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),rep("US",3), "Europe", rep("Japan", 3), rep("US",4), rep("Europe", 3), "US", rep("Europe", 3))
mtcars.country
ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country)

ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country)
```


Graphical parameters with ggbiplot. Using circle and scal. Then try to remove arrow by var.axes = FALSE. Customize ggbiplot.
Specify the colours to use for the groups with scale_colour_manual(). Add a title with ggtitle(). Specify the minimal() theme. Move the legend with theme().
```{r, echo=FALSE}
#Graphical parameters with ggbiplot
##Use circle
ggbiplot(mtcars.pca,ellipse=TRUE,circle=TRUE, labels=rownames(mtcars), groups=mtcars.country)

##Use scale
ggbiplot(mtcars.pca,ellipse=TRUE,obs.scale = 1, var.scale = 1,  labels=rownames(mtcars), groups=mtcars.country)

##Remove arrow var.axes=FALSE
ggbiplot(mtcars.pca,ellipse=TRUE,obs.scale = 1, var.scale = 1,var.axes=FALSE,   labels=rownames(mtcars), groups=mtcars.country)

#Customize ggbiplot
#Specify the colours to use for the groups with scale_colour_manual()

#Add a title with ggtitle()

#Specify the minimal() theme

#Move the legend with theme()
ggbiplot(mtcars.pca,ellipse=TRUE,obs.scale = 1, var.scale = 1,  labels=rownames(mtcars), groups=mtcars.country) +
  scale_colour_manual(name="Origin", values= c("forest green", "red3", "dark blue"))+
  ggtitle("PCA of mtcars dataset")+
  theme_minimal()+
  theme(legend.position = "bottom")
```


Adding a new sample to compare resutls to the original analysis.
Adding a car called spacecar. country in Jupiter. Project a new sample onto the original PCA. Ploting the PCA result to the original data.
```{r}
#Adding a new sample
##adding a car called spacecar. country in Jupiter
spacecar <- c(1000,60,50,500,0,0.5,2.5,0,1,0,0)

mtcarsplus <- rbind(mtcars, spacecar)
mtcars.countryplus <- c(mtcars.country, "Jupiter")

mtcarsplus.pca <- prcomp(mtcarsplus[,c(1:7,10,11)], center = TRUE,scale. = TRUE)

ggbiplot(mtcarsplus.pca, obs.scale = 1, var.scale = 1, ellipse = TRUE, circle = FALSE, var.axes=TRUE, labels=c(rownames(mtcars), "spacecar"), groups=mtcars.countryplus)+
  scale_colour_manual(name="Origin", values= c("forest green", "red3", "violet", "dark blue"))+
  ggtitle("PCA of mtcars dataset, with extra sample added")+
  theme_minimal()+
  theme(legend.position = "bottom")

#Project a new sample onto the original PCA
s.sc <- scale(t(spacecar[c(1:7,10,11)]), center= mtcars.pca$center)
s.pred <- s.sc %*% mtcars.pca$rotation


mtcars.plusproj.pca <- mtcars.pca
mtcars.plusproj.pca$x <- rbind(mtcars.plusproj.pca$x, s.pred)


ggbiplot(mtcars.plusproj.pca, obs.scale = 1, var.scale = 1, ellipse = TRUE, circle = FALSE, var.axes=TRUE, labels=c(rownames(mtcars), "spacecar"), groups=mtcars.countryplus)+
  scale_colour_manual(name="Origin", values= c("forest green", "red3", "violet", "dark blue"))+
  ggtitle("PCA of mtcars dataset, with extra sample projected")+
  theme_minimal()+
  theme(legend.position = "bottom")
```

